Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-12-03 Cost_manuscript.Rmd
reframing the modeling section

@@

  such that a trial’s reward $R_i$ should be accepted if it is larger than the average amount one could earn per second in the environment $\gamma$ (gamma; the OC) during the handling time $H_i$. Individuals with higher gammas therefore required higher rewards in order to accept a given offer. We fit this and other models using a softmax function to predict the probability of accepting a given trial

@@

  All free paremeters of the softmax function (in this case, $\gamma$ and the temperature parameter $\beta$) were estimated independently for each participant, using a maximum likelihood (MLE) approach through the NLOPTR package (CITE **GARY INFO**). The gamma optimization search space was bound by extreme choice values. For example, someone with an extremely high OC would not accept even the most beneficial offers, and would reject a 20 cent offer when the handling time was 2 seconds (or 20/2). Conversely, a participant experiencing low opportunity costs would accept the non-beneficial offer of 4 cents for a 14 second delay (or 4/14). **We cross-validated each subject’s OC value using the pre-midpoint data for estimation, and post-midpoint choices for testing. The estimates from the training sample were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each group.** We then used a two-sided one-sample t-tests to compare each group's gammas to the average optimal earning rate (0.61, see Operationalization of Cost), allowing us to determine whether any demand boosted an offer's value. Finally, we compared gammas with an ANOVA (with condition as a factor), followed by post-hoc permutations (5000 iterations) comparing mean gammas per group.

  The base formulation lacked the flexibility needed to capture the dynamics observed in the within-subjects experiment. We therefore tested a second model which allowed $\gamma$ to evolve over time, adapting a learning rule from @Constantino2015

@@

  such that the handling time was counted only for accepted trials. Gamma was initialized by the median earning rate achieved across individuals. These modifications allowed for the tracking of recent history that was part of our criteria, while mostly maintaining the same complexity (i.e. one free parameter) and structure of the base model

@@

  A final model considered the relationship between effortful engagement and the perceived length of time. Being immersed in a task can make time transpire subjectively faster [@Csikszentmihalyi2014], which might have made the cognitive task attractive. Time sensitivity has been previously used to improve fits on classic hyperbolic models of delay discounting [@Mckerchar2009]. It is possible that such subjective handling time was recalibrated in the within-subjects experiment, resulting in the observed convergance of acceptances across demands. We therefore added a time sensitivity parameter $S$ to the model, with a separate value associated with each type of demand. The values of $S_{demand}$ could then be recalibrated as participants experienced them, averaged according to the following formula [adapted from @Sutton2018]

@@

  In order to compare model fits, we used expectation maximization...(**Gary**)

@@

>>>>

  **GARY. All free paremeters of the softmax function (the temperature parameter $\beta$) were estimated independently for each participant, using a maximum likelihood (MLE) approach through the NLOPTR package. The gamma optimization search space was bound by extreme choice values. For example, someone with an extremely high OC would not accept even the most beneficial offers, and would reject a 20 cent offer when the handling time was 2 seconds (or 20/2). Conversely, a participant experiencing low opportunity costs would accept the non-beneficial offer of 4 cents for a 14 second delay (or 4/14). We cross-validated each subject’s OC value using the pre-midpoint data for estimation, and post-midpoint choices for testing. The estimates from the training sample were used to predict acceptances in the testing sample, and the mean percent correctly predicted was reported for each group.**

@@

  such that a trial’s reward $R_i$ should be accepted if it is larger than the average amount one could earn per second in the environment $\gamma$ (gamma; the OC) during the handling time $H_i$, with high gammas promoting selectivity. We fit this and all subsequent models using a softmax function to predict the probability of accepting a given trial

@@

  gamma in this base model was computed as an individual's empirical earning rate by the time of each offer, with the free temperature parameter $\beta$ indexing decision noise. However, this base formulation lacked the flexibility needed to capture the dynamics observed in the within-subjects experiment. We therefore tested a second model which allowed $\gamma$ to evolve over time, adapting a learning rule from @Constantino2015

@@

  such that the handling time was counted only for accepted trials. For this adaptive model, we fit the learning rate $\alpha$ and the initial value of $\gamma$ in addition to the temperature. These modifications allowed for the tracking of recent history that was part of our criteria, while maintaining the general structure of the base model

@@

  A third model considered the possibility that each type of demand biased the estimate of $\gamma$. Such bias would be stable if a single demand was experienced, but recalibrated into a global estimate as decision makers faced different types of demands. This dynamic could underlie the observed convergance of acceptances across demands. At each decision point, the bias produced by each demand ($c_{demand}$) was weighted against a global bias estimate $C$

$$\begin{aligned}
mC_i = c_{demand} w_i + C (1 - w_i)
\end{aligned}$$

  We therefore added a time sensitivity parameter $S$ to the model, with a separate value associated with each type of demand. The values of $S_{demand}$ could then be recalibrated as participants experienced them, averaged according to the following formula [adapted from @Sutton2018]

@@

  *(LATER) We then used a two-sided one-sample t-tests to compare each group's gammas to the average optimal earning rate (0.61, see Operationalization of Cost), allowing us to determine whether any demand boosted an offer's value. Finally, we compared gammas with an ANOVA (with condition as a factor), followed by post-hoc permutations (5000 iterations) comparing mean gammas per group.*

@@

- Being immersed in a task can make time transpire subjectively faster [@Csikszentmihalyi2014], which might have made the cognitive task attractive. Time sensitivity has been previously used to improve fits on classic hyperbolic models of delay discounting [@Mckerchar2009].
